### 2019-01-17 Sections/economics_of_the_quantum_internet.tex
Improving payoff enhancement

The time-derivative of the cost of computation is strictly negative, assuming correctness of the growth and cost postulates (Pos.~\ref{post:net_growth} \& \ref{post:cost_comp}),
\begin{align}
\frac{\partial L}{\partial t} \leq 0,	
\end{align}
which implies monotonic reduction in the cost of computation over time, unless network growth and cost completely freeze (\mbox{$\gamma_N=\gamma_C=0$}), in which case the cost of computation remains flat.

@@

Additionally, we have assumed no compute-cycles are wasted -- compute-time is always fully utilised, as per Pos.~\ref{post:market_eff}. Then it follows that the time-share of the combined resources of the entire network should be unity,
\begin{align}
	r_{n_\mathrm{global}}=1.
\end{align}
\mbox{$r_{n_\mathrm{global}}<1$} would imply inefficiency via wasted clock-cycles. Combining this with the linearity constraint implies the arbitrage-free time-sharing model,
\begin{definition}[Arbitrage-free time-sharing model] \label{def:arb_free_ts}\index{Arbitrage-free!Time-sharing model}
In an efficient market for unified quantum computing time-shares\index{Time-sharing}, a network participant in possession of $n$ of the entire $n_\mathrm{global}$ qubits in the network is entitled to the fraction of unified network compute time,
\begin{align}\index{Time-shared compute-time}
	r_n = \frac{n}{n_\mathrm{global}},
\end{align}
where,
\begin{align}
n_\mathrm{global} = \sum_{j\in\mathrm{nodes}} n_j,
\end{align}
is the total number of qubits in the network, and,
\begin{align}
0\leq r_n \leq 1.	
\end{align}
\mbox{$r_n=1$} iff the node has a complete monopoly over qubits, i.e \mbox{$n=n_\mathrm{global}$}.
\end{definition}

@@

Using our postulate for network growth (Pos.~\ref{post:net_growth}) yields the postulated time-dependent QCL,
\begin{postulate}[Time-dependent QCL]
The time-dependent QCL, based on the postulate of exponential network growth, is,
\begin{align}\index{Time-dependent quantum computational leverage}
\lambda_n(t) &= \frac{\chi_\mathrm{sc}(N_0{\gamma_N}^t)}{\chi_\mathrm{sc}(n)},\nonumber\\
\lambda_n^\mathrm{dB}(t) &= 10\log_{10}(\lambda_n(t)).
\end{align}
The initial (\mbox{$t=0$}) time-dependent QCL reduces to the standard QCL formula.
\end{postulate}
Note that for any super-linear scaling function, the time-dependent QCL grows exponentially over time, unlike the classical case where there is no leverage, which does not change over time (i.e \mbox{$\lambda_n(t)=1\,\,\forall\,n,t$}).

@@

As discussed in relation to combined computational scaling functions (Sec.~\ref{sec:comb_comp_sc_func})\index{Combined computational scaling functions}, different market participants will be executing different software applications on their share of the quantum computing resources, with differing QCLs\index{Quantum computational leverage}. Because the applications differ between users, as do their QCLs, so too does the monetary value of the computations they are performing. This yields the distinction between \textit{subjective}\index{Subjective value of computation} and \textit{objective}\index{Objective value of computation} value of computation:

@@

Now, quantum computing uses similar strategies for possible solutions, with one key difference. Using classical computing, the relationship for the \textsc{Unify} strategy was described as additive in nature, where the computational resources of both players are accumulated additively when unified into a larger virtual computer. In the corresponding quantum \textsc{Unify} strategy, this effect is enhanced by their super-linear computational leverage. For the sake of illustration, we will now assume this becomes multiplicative. As such, should $X$ and $Y$ build identical quantum computers, the hypothesised payoff matrix would become as shown in Tab.~\ref{tab:game_theory_2}.

@@

When implementing a quantum computation characterised by completely general computational scaling functions\index{Computational scaling functions}, $f^{X,Y}_\mathrm{sc}$ (which in general can be distinct for the two players $X$ and $Y$), elements in the payoff matrix associated with unification will be enhanced to \mbox{$(c_X,c_Y)$}, where,
\begin{align}
c_X &= \frac{n_X\cdot f^X_\mathrm{sc}(n_X+n_Y)}{n_X+n_Y},\nonumber\\
c_Y &= \frac{n_Y\cdot f^Y_\mathrm{sc}(n_X+n_Y)}{n_X+n_Y},
\end{align}
where the payoff matrix elements are assumed to be in units of classical-equivalent processing time, and resource sharing is based on the techniques for arbitrage-free time-share allocation presented in Sec.~\ref{sec:arb_free_time_share}.

Thus, it is evident that the enhancement in cooperative payoff matrix elements is highly dependent on:
\begin{itemize}
	\item The total number of qubits held by the two players.
	\item The asymmetry in the distribution of qubits between them.
	\item The respective algorithms to which the computational resources are being applied by each player.
\end{itemize}

The final point is particularly noteworthy, since it implies that optimal game-theoretic outcomes are not objective, but subjective, and highly dependent on how players are employing their computational resources, which may be highly asymmetric. If one player is employing an exponential scaling function, whereas the other is only employing a polynomial one, this could completely distort the utility payoff dynamics of the game in favour of the player who would otherwise have been weaker under symmetric scaling functions. This in turn could completely alter the landscape of how users choose strategies to play optimally.

>>>>

The time-derivative of the cost of computation is strictly negative, assuming correctness of the growth and cost postulates (Post.~\ref{post:net_growth} \& \ref{post:cost_comp}),
\begin{align}
\frac{\partial L}{\partial t} \leq 0,	
\end{align}
which implies monotonic reduction in the cost of computation over time, unless network growth and cost completely freeze (\mbox{$\gamma_N=\gamma_C=0$}), in which case the cost of computation remains flat.

@@

Additionally, we have assumed no compute-cycles are wasted -- compute-time is always fully utilised, as per Post.~\ref{post:market_eff}. Then it follows that the time-share of the combined resources of the entire network should be unity,
\begin{align}
	r_{n_\mathrm{global}}=1.
\end{align}
\mbox{$r_{n_\mathrm{global}}<1$} would imply inefficiency via wasted clock-cycles. Combining this with the linearity constraint implies the arbitrage-free time-sharing model,
\begin{definition}[Arbitrage-free time-sharing model] \label{def:arb_free_ts}\index{Arbitrage-free!Time-sharing model}
In an efficient market for unified quantum computing time-shares\index{Time-sharing}, a network participant in possession of $n$ of the entire $n_\mathrm{global}$ qubits in the network is entitled to the fraction of unified network compute time,
\begin{align}\index{Time-shared compute-time}
	r_n = \frac{n}{n_\mathrm{global}},
\end{align}
where,
\begin{align}
n_\mathrm{global} = \sum_{j\in\mathrm{nodes}} n_j,
\end{align}
is the total number of qubits in the network, and,
\begin{align}
0\leq r_n \leq 1.	
\end{align}
\mbox{$r_n=1$} iff the node has a complete monopoly over qubits, i.e \mbox{$n=n_\mathrm{global}$}.
\end{definition}

@@

Using our postulate for network growth (Post.~\ref{post:net_growth}) yields the postulated time-dependent QCL,
\begin{postulate}[Time-dependent QCL]
The time-dependent QCL, based on the postulate of exponential network growth, is,
\begin{align}\index{Time-dependent quantum computational leverage}
\lambda_n(t) &= \frac{\chi_\mathrm{sc}(N_0{\gamma_N}^t)}{\chi_\mathrm{sc}(n)},\nonumber\\
\lambda_n^\mathrm{dB}(t) &= 10\log_{10}(\lambda_n(t)).
\end{align}
The initial (\mbox{$t=0$}) time-dependent QCL reduces to the standard QCL formula.
\end{postulate}
Note that for any super-linear scaling function, the time-dependent QCL grows exponentially over time, unlike the classical case where there is no leverage, which does not change over time (i.e \mbox{$\lambda_n(t)=1\,\,\forall\,n,t$}).

@@

As discussed in relation to combined computational scaling functions (Sec.~\ref{sec:comb_comp_sc_func})\index{Combined computational scaling functions}, different market participants will be executing different software applications on their share of the quantum computing resources, with differing QCLs\index{Quantum computational leverage}. Because the applications differ between users, as do their computational scaling functions ($f_\mathrm{sc}$), QCLs, so too does the monetary value of the computations they are performing. This yields the distinction between \textit{subjective}\index{Subjective value of computation} and \textit{objective}\index{Objective value of computation} value of computation:

@@

Now, quantum computing uses similar strategies for possible solutions, with one key difference. Using classical computing, the relationship for the \textsc{Unify} strategy was described as additive in nature, where the computational resources of both players are accumulated additively when unified into a larger virtual computer. In the corresponding quantum \textsc{Unify} strategy, this effect is enhanced by their super-linear computational leverage. For the sake of illustration, we will now assume this becomes multiplicative. Multiplicativity in computational power approximates behaviour under a \textsc{Unify} strategy when dealing with exponential scaling functions.

This is easily intuitively seen as follows. Let the computational scaling function be an arbitrary exponential,
\begin{align}
f_\mathrm{sc}(n) = O(\mathrm{exp}(n)).	
\end{align}
Then it immediately follows that the scaling function obeys the identity,
\begin{align}
	f_\mathrm{sc}\left(\sum_i n_i \right) = \prod_i f_\mathrm{sc}(n_i),
\end{align}
yielding the multiplicative behaviour.

As such, should $X$ and $Y$ build identical quantum computers, the hypothesised payoff matrix would become as shown in Tab.~\ref{tab:game_theory_2}.

@@

When implementing a quantum computation characterised by completely general computational scaling functions\index{Computational scaling functions}, $f^{X,Y}_\mathrm{sc}$ (which in general can be distinct for the two players, $X$ and $Y$), elements in the payoff matrix associated with unification will undergo the quantum utility payoff enhancement,
\begin{align}\label{eq:util_payoff_enh_1}
(f^X_\mathrm{sc}(n_X),f^Y_\mathrm{sc}(n_Y))	\underset{\textsc{Unify}}{\longrightarrow} (c_X,c_Y),
\end{align}
where,
\begin{align}\label{eq:util_payoff_enh_2}
c_X &= \frac{n_X\cdot f^X_\mathrm{sc}(n_X+n_Y)}{n_X+n_Y},\nonumber\\
c_Y &= \frac{n_Y\cdot f^Y_\mathrm{sc}(n_X+n_Y)}{n_X+n_Y},
\end{align}
and payoff matrix elements are assumed to be in units of classical-equivalent processing time\footnote{There's nothing unique or special about using classical-equivalent computational power as our utility measure. Any other measure of `payoff' could be equally well justified, depending on circumstance. For example, one could instead represent utility in terms of the monetary cost of computational power. In that case, we simply need to transform the matrix elements in the payoff matrix using the cost of computation\index{Cost of computation} identity presented in Post.~\ref{post:cost_comp}.} (i.e FLOPs), and resource-sharing is based on the methodology for arbitrage-free time-share allocation presented in Sec.~\ref{sec:arb_free_time_share}.

Thus, it is evident that the enhancement in cooperative payoff matrix elements is highly dependent on:
\begin{itemize}
	\item The total number of qubits held between the two players, \mbox{$n_X+n_Y$}.
	\item The asymmetry in the distribution of qubits between them, \mbox{$n_X/n_Y$}.
	\item The respective algorithms to which the computational resources are being applied by each player, which influence the subjective scaling functions $f^X_\mathrm{sc}$ and $f^Y_\mathrm{sc}$ independently.
\end{itemize}

The final point is particularly noteworthy, since it implies that optimal game-theoretic outcomes are not objective, but subjective, and highly dependent on how players are employing their computational resources, which may be highly asymmetric and change dynamically over time. If one player is employing an exponential scaling function, whereas the other is only employing a polynomial one, this could completely distort the utility payoff dynamics of the game in favour of the player who would otherwise have been weaker under symmetric scaling functions. This in turn could completely alter the landscape of how users choose strategies to play optimally.

The utility payoff enhancement characterised by the transformation given in Eqs.~(\ref{eq:util_payoff_enh_1} \& \ref{eq:util_payoff_enh_2}) is the simplest of toy models where unification is assumed to be perfectly efficient -- there no overheads associated with unification, nor are there any externalities, such as taxes or regulations. However, such secondary effects can easily be incorporated into the model by modulating Eq.~(\ref{eq:util_payoff_enh_2}) with additional layers of transformations capturing these features. For example, consider the unification of computational resources between two players residing in different jurisdictions, which levy import/export tariffs\index{Tariffs} against one another, an externality introducing inefficiency into cooperative strategies. When expressed in terms of the monetary cost of computation, this would effectively modulate the payoffs of \textsc{Unify} matrix elements by a tariff-dependent function.
